---
title: |
  | Black and Latinxs Students in Political Science
  | NCES IPED Data 2011-2017
author: "Clara Bicalho"
date: "6/12/2020"
output: pdf_document
---

The descriptive data shown below was taken from the [National Center for Eduation Statistics (NCES)](https://nces.ed.gov/ipeds/). The only data set that contained race and ethinicity counts by year were that of completed degrees. NCES/IPEDS Completion data do not include race/ethnicity of non-resident aliens. These are therefore excluded from the group-based counts, though these students are included in the denominator in the proportion column. Top 15 programs in Tables 1-2 are according to [USNews 2017 Rankings](https://www.usnews.com/best-graduate-schools/top-humanities-schools/political-science-rankings).

<!-- Figure 1 shows the share of Black and Latinxs students who earned Political Science degrees between 2011-2017. The peaks in Doctorate Degrees awarded to Black students were all years in which 1 student was awarded a degree. Similarly, there was 1 PhD degree awarded for a Latinx student in 2011, 2 in 2013, and 1 in 2018. Figure 2 shows the same statistics for the Sociology Department, for comparison. -->

More details on data base [here](https://surveys.nces.ed.gov/ipeds/VisInstructions.aspx?survey=10&id=30080&show=all).

\clearpage

\centering

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(xtable)
library(dplyr)
from_scratch <- FALSE
```

```{r import, message=FALSE, warning=FALSE}
if(from_scratch){
  files <- paste0("data/C", 2011:2018, "_A.csv")
dat <- do.call(plyr::rbind.fill, lapply(files, function(f){
  df <- read_csv(f)
  df$Year <- substr(f, 7, 10)
  return(df)
  }))
save(dat, "completion_2011-2018.RDS")
}else{
 dat <- readRDS("completion_2011-2018.RDS")
}


ps <- dat %>% 
  filter(
    CIPCODE %in% c("45.1099", "45.1001", "45.1002", "45.1003", "45.1004") & 
      UNITID %in% c(110635, 186131, 130794, 193900, 166027, 166683, 243744,
                    110680, 110662, 170976, 199120, 190150, 198419, 144050,
                    204796)) %>%
  mutate(University = case_when(UNITID == 110635 ~ "UC Berkeley",
                                UNITID == 186131 ~ "Princeton",
                                UNITID == 130794 ~ "Yale",
                                UNITID == 193900 ~ "NYU",
                                UNITID == 166027 ~ "Harvard",
                                UNITID == 166683 ~ "MIT",
                                UNITID == 243744 ~ "Stanford",
                                UNITID == 110680 ~ "UCSD",
                                UNITID == 110662 ~ "UCLA",
                                UNITID == 170976 ~ "UMich",
                                UNITID == 199120 ~ "UNC-Chapel Hill",
                                UNITID == 190150 ~ "Columbia",
                                UNITID == 198419 ~ "Duke",
                                UNITID == 144050 ~ "U Chicago",
                                UNITID == 204796 ~ "Ohio State University"))
```

```{r summary}
tab1 <- ps %>% group_by(University) %>% 
  summarize(`Total Black` = sum(CBKAAT[AWLEVEL == 17], na.rm=TRUE),
            `Share Black` = sum(CBKAAT[AWLEVEL == 17], na.rm=TRUE)/
              sum(CTOTALT[AWLEVEL == 17], na.rm=TRUE),
            `Total Latinxs` = sum(CHISPT[AWLEVEL == 17], na.rm=TRUE),
            `Share Latinxs` = sum(CHISPT[AWLEVEL == 17], na.rm=TRUE)/
              sum(CTOTALT[AWLEVEL == 17], na.rm=TRUE))
```

```{r, include=FALSE}
# Undegrads
tab2 <- ps %>% group_by(University) %>% 
  summarize(`Total Black` = sum(CBKAAT[AWLEVEL == "05"], na.rm=TRUE),
            `Share Black` = sum(CBKAAT[AWLEVEL == "05"], na.rm=TRUE)/
              sum(CTOTALT[AWLEVEL == "05"], na.rm=TRUE),
            `Total Latinxs` = sum(CHISPT[AWLEVEL == "05"], na.rm=TRUE),
            `Share Latinxs` = sum(CHISPT[AWLEVEL == "05"], na.rm=TRUE)/
              sum(CTOTALT[AWLEVEL == "05"], na.rm=TRUE))

# ps %>% select(CBKAAT, UNITID)
table(ps$AWLEVEL)
```

```{r export, echo=FALSE, results="asis"}
exp_tab1 <- capture.output(print(
  xtable(tab1, caption = "Black and Latinxs Students Who Completed a Political Science PhD Program, Top 15 Programs (2011-2017)"), align = "tbph",
  caption.placement = "top", include.rownames = FALSE, comment = FALSE,
  header = FALSE))

cat(exp_tab1, sep = "\n")
```

```{r export2, echo=FALSE, results="asis"}
exp_tab2 <- capture.output(print(
  xtable(tab2, caption = "Black and Latinxs Students Who Completed a Political Science B.A., Top 15 Programs (2011-2017)"), align = "tbph",
  caption.placement = "top", include.rownames = FALSE, comment = FALSE,
  header = FALSE))

cat(exp_tab2, sep = "\n")
```